---
title: "Analysis"
author: "Kiran Smelser"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(dplyr)
library(tidyverse)
library(data.table)
library(readxl)
library(mltools)
library(ggstatsplot)
```

```{r}
path_data="./data/Supplementary_Table - Study Data by Groups.csv"
path_classifier="./data/ciitizen_health_classifier.xlsx"
```

```{r}
data <- read_csv(path_data)
```


#1 Correlations Between Groups and Seizure Types
##1.1 Initial Seizure Type
```{r}
type_data <- data
new_col_names <- gsub("seizure_history_type_", "", colnames(type_data))
colnames(type_data) <- new_col_names

data_long <- type_data %>%
  gather(key = "initial_seizure_type", value = "value1", focal:infantile) %>%
  gather(key = "seizure_type", value = "value2", `Absence seizure`:`Unprovoked seizure`) %>%
  filter(value1 == 1, value2 == 1) %>%
  group_by(seizure_type) %>%
  mutate(n = n()) %>%
  filter(n >= 15)

# Plot
p_init <- ggbarstats(data_long, initial_seizure_type, seizure_type) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##1.2 Number of Seizure Types
```{r}
data_long <- type_data %>%
  gather(key = "seizure_type", value = "value2", `Absence seizure`:`Unprovoked seizure`) %>%
  filter(value2 == 1) %>%
  group_by(seizure_type) %>%
  mutate(n = n()) %>%
  filter(n >= 5)

# Plot
p_unique <- ggbarstats(data_long, unique_types, seizure_type) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##1.3 Abnormal EEG
```{r}
data_long <- data_long %>%
  filter(n >= 7)

# Plot
p_abnormal <- ggbarstats(data_long, abnormal_eeg, seizure_type) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##1.4 Age of Onset
```{r}
data_long <- data_long %>%
  filter(n >= 10)

# Plot
p_onset <- ggbarstats(data_long, Onset_group, seizure_type) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##1.5 Save Plots
```{r}
plot <- list(p_init, p_unique, p_abnormal, p_onset)
pdf("./figures/type_plots.pdf", width = 12, height = 8)
plot
dev.off()
```


#2 Correlations Between Groups and Hospitalization Events
##2.1 Initial Seizure Type
```{r}
hospital_data <- data
new_col_names <- gsub("admission_diagnosis_", "", colnames(hospital_data))
colnames(hospital_data) <- new_col_names

data_long <- hospital_data %>%
  gather(key = "initial_seizure_type", value = "value1", focal:infantile) %>%
  gather(key = "hospital_event", value = "value2", `Abnormal movement`:Vomiting) %>%
  filter(value1 == 1, value2 == 1) %>%
  group_by(hospital_event) %>%
  mutate(n = n()) %>%
  filter(n >= 5)

# Plot
p_init <- ggbarstats(data_long, initial_seizure_type, hospital_event) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##2.2 Number of Seizure Types
```{r}
data_long <- hospital_data %>%
  gather(key = "hospital_event", value = "value2", `Abnormal movement`:Vomiting) %>%
  filter(value2 == 1) %>%
  group_by(hospital_event) %>%
  mutate(n = n()) %>%
  filter(n >= 5)

# Plot
p_unique <- ggbarstats(data_long, unique_types, hospital_event) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##2.3 Abnormal EEG
```{r}
data_long <- data_long %>%
  filter(n >= 7)

# Plot
p_abnormal <- ggbarstats(data_long, abnormal_eeg, hospital_event) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##2.4 Age of Onset
```{r}
# Plot
p_onset <- ggbarstats(data_long, Onset_group, hospital_event) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##2.5 Save Plots
```{r}
plot <- list(p_init, p_unique, p_abnormal, p_onset)
pdf("./figures/hospital_plots.pdf", width = 12, height = 8)
plot
dev.off()
```


#3 Correlations Between Groups and Clinical Diagnoses
##3.1 Initial Seizure Type
```{r}
clinical_data <- data
new_col_names <- gsub("clinical_diagnosis_", "", colnames(clinical_data))
colnames(clinical_data) <- new_col_names

data_long <- clinical_data %>%
  gather(key = "initial_seizure_type", value = "value1", focal:infantile) %>%
  gather(key = "clinical_diagnosis", value = "value2", `Abdominal pain`:`West syndrome`) %>%
  filter(value1 == 1, value2 == 1) %>%
  group_by(clinical_diagnosis) %>%
  mutate(n = n()) %>%
  filter(n > 30)

# Plot
p_init <- ggbarstats(data_long, initial_seizure_type, clinical_diagnosis) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##3.2 Number of Seizure Types
```{r}
data_long <- clinical_data %>%
  gather(key = "clinical_diagnosis", value = "value2", `Abdominal pain`:`West syndrome`) %>%
  filter(value2 == 1) %>%
  group_by(clinical_diagnosis) %>%
  mutate(n = n()) %>%
  filter(n >= 15)

# Plot
p_unique <- ggbarstats(data_long, unique_types, clinical_diagnosis) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##3.3 Abnormal EEG
```{r}
# Plot
p_abnormal <- ggbarstats(data_long, abnormal_eeg, clinical_diagnosis) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##3.4 Age of Onset
```{r}
# Plot
p_onset <- ggbarstats(data_long, Onset_group, clinical_diagnosis) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##3.5 Save Plots
```{r}
plot <- list(p_init, p_unique, p_abnormal, p_onset)
pdf("./figures/clinical_plots.pdf", width = 12, height = 8)
plot
dev.off()
```


#4 Correlations Between Groups and Diagnostic Procedure Results
##4.1 Initial Seizure Type
```{r}
procedure_data <- data
new_col_names <- gsub("procedure_findings_", "", colnames(procedure_data))
colnames(procedure_data) <- new_col_names

data_long <- procedure_data %>%
  gather(key = "initial_seizure_type", value = "value1", focal:infantile) %>%
  gather(key = "procedure_result", value = "value2", `Abnormal restricted diffusion`:`Videofluoroscopy swallow normal`) %>%
  filter(value1 == 1, value2 == 1) %>%
  group_by(procedure_result) %>%
  mutate(n = n()) %>%
  filter(n >= 20)

# Plot
p_init <- ggbarstats(data_long, initial_seizure_type, procedure_result) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##4.2 Number of Seizure Types
```{r}
data_long <- procedure_data %>%
  gather(key = "procedure_result", value = "value2", `Abnormal restricted diffusion`:`Videofluoroscopy swallow normal`) %>%
  filter(value2 == 1) %>%
  group_by(procedure_result) %>%
  mutate(n = n()) %>%
  filter(n >= 10)

# Plot
p_unique <- ggbarstats(data_long, unique_types, procedure_result) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##4.3 Abnormal EEG
```{r}
# Plot
p_abnormal <- ggbarstats(data_long, abnormal_eeg, procedure_result) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##4.4 Age of Onset
```{r}
# Plot
p_onset <- ggbarstats(data_long, Onset_group, procedure_result) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##4.5 Save Plots
```{r}
plot <- list(p_init, p_unique, p_abnormal, p_onset)
pdf("./figures/procedure_plots.pdf", width = 12, height = 8)
plot
dev.off()
```


#5 Correlations Between Groups and Number of Stable Medications
##5.1 Initial Seizure Type
```{r}
stable_data <- data

data_long <- stable_data %>%
  gather(key = "initial_seizure_type", value = "value1", focal:infantile) %>%
  filter(value1 == 1)

# Plot
p_init <- ggbarstats(data_long, initial_seizure_type, num_stable_meds)
```

##5.2 Number of Seizure Types
```{r}
# Plot
p_unique <- ggbarstats(stable_data, unique_types, num_stable_meds)
```

##5.3 Abnormal EEG
```{r}
# Plot
p_abnormal <- ggbarstats(stable_data, abnormal_eeg, num_stable_meds) 
```

##5.4 Age of Onset
```{r}
# Plot
p_onset <- ggbarstats(stable_data, Onset_group, num_stable_meds) 
```

##5.5 Save Plots
```{r}
plot <- list(p_init, p_unique, p_abnormal, p_onset)
pdf("./figures/stable_plots.pdf", width = 12, height = 8)
plot
dev.off()
```