---
title: "Grouping Patients"
author: "Kiran Smelser"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(dplyr)
library(tidyverse)
library(data.table)
library(readxl)
library(mltools)
```

```{r}
path_data="./data/Ciitizen_SCN8A_UArizona_2024.02.09.xlsx"
path_classifier="./data/ciitizen_health_classifier.xlsx"
```


#1. Assigning Groups
```{r}
Ciitizen_variant = read_excel(path_data, sheet = "genetic_findings")
scn8a_variant <- Ciitizen_variant %>% filter(gene == "SCN8A")

scn8a_variant <- scn8a_variant %>% group_by(patient_uuid) %>% rename("UUID" = "patient_uuid")

scn8a_variant <- scn8a_variant %>% distinct(UUID, .keep_all = T)

scn8a_variant <- scn8a_variant %>% select(c(UUID, variant_DNA, variant_protein))
```

##1.1 Age of Onset
```{r}
onset_age <- read.csv("./data/onset_ages.csv")

onset_age <- data.table(onset_age)
onset_age <- onset_age[, min(Age_onset), by = "UUID"]

onset_age <- onset_age %>% mutate( Onset_group = 
                                     ifelse(V1 <=30, "Neonatal",
                                            ifelse(V1 >30 & V1 <= 120, "1-4 Months",
                                                   ifelse(V1 > 120 & V1 <= 210, "4-7 Months",
                                                          ifelse(V1 > 210 & V1 <=365, "7-12 Months", "12+ Months")))))
study <- onset_age %>% rename("age_onset" = "V1")
study <- study %>% mutate(age_onset_m = age_onset/30)
study <- left_join(scn8a_variant, study, by = "UUID")
```


##1.2 Initial Seizure Type - Focal, Bilateral_TC, Absence, Infantile_spasms
```{r}
# Import inital seizure type classifier
initial_classifier <- read_excel("./data/inital_seizure_types_classifier.xlsx")


# Import Seizure data
Ciitizen_seizure = read_excel(path_data, sheet = "seizure_history")
Ciitizen_seizure <- Ciitizen_seizure[, c("patient_uuid",	"seizure_history_type", "seizure_history_age_days")]


# Filter for range 0-3 years old
Ciitizen_seizure <- Ciitizen_seizure %>% filter(seizure_history_age_days < 1095)

# One-hot encode Focal, Bilateral_TC, Absence, Infantile_spasms seizure
Ciitizen_seizure$focal <- ifelse(Ciitizen_seizure$seizure_history_type %in% initial_classifier$focal, 1, 0)
Ciitizen_seizure$bilateral_tc <- ifelse(Ciitizen_seizure$seizure_history_type %in% initial_classifier$bilateral_tonic_clonic, 1, 0)
Ciitizen_seizure$absence <- ifelse(Ciitizen_seizure$seizure_history_type %in% initial_classifier$absence, 1, 0)
Ciitizen_seizure$infantile <- ifelse(Ciitizen_seizure$seizure_history_type %in% initial_classifier$infantile_spasms, 1, 0)

Ciitizen_seizure <- Ciitizen_seizure %>% rename("UUID" = "patient_uuid")
initial_seizures <- Ciitizen_seizure[, c("UUID", "focal", "bilateral_tc", "absence", "infantile")]
initial_seizures <- initial_seizures %>%
  group_by(UUID) %>%
  summarise(across(everything(), sum, na.rm = TRUE))
initial_seizures <- initial_seizures %>%
  group_by(UUID) %>%
  mutate(across(everything(), ~ifelse(. > 0, 1, .)))
study <- left_join(study, unique(initial_seizures), by = "UUID")
```

##1.3 Abnormal EEGs
```{r}
# Import classifier
classifier <- read_excel(path_classifier)

# Import EEG data
Ciitizen_eegs = read_excel(path_data, sheet = "diagnostic_procedures")
Ciitizen_eegs <- Ciitizen_eegs[grepl("EEG", Ciitizen_eegs$procedure),]
Ciitizen_eegs <- subset(Ciitizen_eegs, select = -c(2:5, 8:9))

# Filter for range 0-3 years old
Ciitizen_eegs <- Ciitizen_eegs %>% filter(procedure_age_days < 1095)

# One-hot encode abnormal EEG status
Ciitizen_eegs$abnormal_eeg <- with(Ciitizen_eegs, ave(!(procedure_findings %in% classifier$eeg_normal), patient_uuid, FUN = function(x) as.integer(any(x))))
Ciitizen_eegs <- Ciitizen_eegs %>% rename("UUID" = "patient_uuid")
Ciitizen_eegs <- unique(subset(Ciitizen_eegs, select = c(1, 4)))
study <- left_join(study, Ciitizen_eegs, by = "UUID")
```

##1.4 # of Seizure Types
```{r}
# Determine # of Seizure types in the first 3 years
types <- Ciitizen_seizure[, c("UUID", "seizure_history_type")]
types <- types[!types$seizure_history_type %in% c("Seizure", "Unclassified seizure", "Provoked seizure", "Unprovoked seizure"), ]

# Modify classifier
types_classifier <- subset(classifier, select = c(5:10))


# Function to add seizure types
find_column_name <- function(value, df) {
  for (col_name in names(df)) {
    if (value %in% df[[col_name]]) {
      return(col_name)
    }
  }
  return(NA)
}
types$type <- sapply(types$seizure_history_type, find_column_name, df=types_classifier)

types <- types %>%
  group_by(UUID) %>%
  mutate(unique_types = n_distinct(type, na.rm = TRUE))
types <- subset(types, select = -c(2:3))

# Merge
study <- left_join(study, unique(types), by = "UUID")
```


#2. Load more data for building out dataset
```{r}
Ciitizen_diagnosis= read_excel(path_data, sheet = "clinical_diagnosis")
Ciitizen_medication = read_excel(path_data, sheet = "medication_aggregate")
Ciitizen_hospitalization = read_excel(path_data, sheet = "hospital_admission")
Ciitizen_demographics = read_excel(path_data, sheet = "demographics")
Ciitizen_diagnostic = read_excel(path_data, sheet = "diagnostic_procedures")

#only interested in <3 years old here 
Ciitizen_diagnosis <- Ciitizen_diagnosis %>% filter(clinical_diagnosis_age_days_firstDate<1095)
Ciitizen_hospitalization <- Ciitizen_hospitalization %>% filter(admission_age_days_firstDate<1095)
Ciitizen_medication_weaned <- Ciitizen_medication %>% filter(medication_age_days_lastDate < 1095) 
Ciitizen_diagnostic <- Ciitizen_diagnostic %>% filter(procedure_age_days < 1095)
```


#3. Expanding Dataset to include features of interest
##3.1 Getting Medications when stable
```{r}
Ciitizen_medication_stable <- Ciitizen_medication

# Clean aggregated medication data
meds_to_use <- paste(c("Adrenocorticotropin (ACTH 1-18),I-125 (TYR)", "Clonazepam", "Levetiracetam", "Phenytoin", "Oxcarbazepine", "Carbamazepine", "Phenobarbital", "Lamotrigine", "Briveracetam", "Cannabidiol", "Clobazam", "Epidiolex", "Eslicarbazepine", "Ethosuximide", "Felbamate", "Fosphenytoin", "Gabapentin", "Prednisolone", "Lacosamide", "Primidone", "Pyridoxine", "Rufinamide", "Topiramate", "Valproate", "Vigabatrin", "Zonisamide", "Stiripentol", "Tiagabine", "Acetazolamide", "Clonidine", "Perampanel"), collapse = "|")
Ciitizen_medication_stable <- Ciitizen_medication_stable %>%
  mutate(medication = ifelse(grepl("ACTH", medication), "ACTH", medication))
Ciitizen_medication_stable <- subset(Ciitizen_medication_stable, grepl(meds_to_use, medication))
Ciitizen_medication_stable <- subset(Ciitizen_medication_stable, select = -c(3:6, 9:11))
```

```{r}
# Function to find stable periods and count medications
find_stable_periods <- function(data) {
  # Sort by first date and last date
  data <- data[order(data$medication_age_days_firstDate,
                     data$medication_age_days_lastDate),]
  
  max_meds <- 0 
  stable_periods <- NULL
  for (i in seq_len(nrow(data))) {
    start_date <- data$medication_age_days_firstDate[i]
    end_date <- data$medication_age_days_lastDate[i]
    
    overlaps <- subset(data,
                       medication_age_days_firstDate <= start_date &
                         medication_age_days_lastDate >= end_date)
    
    # Check if the period is stable
    if ((end_date - start_date >=365) & (nrow(overlaps) > max_meds)) {
      max_meds <- nrow(overlaps)
      stable_periods <- overlaps
    }
  }
  
  if (is.null(stable_periods)) {
    stable_periods <- data[FALSE,]
  }
  
  result <- data.frame(
    patient_uuid = unique(data$patient_uuid),
    age_days_start = ifelse(is.null(stable_periods), NA, min(stable_periods$medication_age_days_firstDate)),
    age_days_end = ifelse(is.null(stable_periods), NA, max(stable_periods$medication_age_days_lastDate)),
    num_stable_meds = ifelse(is.null(stable_periods), NA, nrow(stable_periods))
  )
  return(result)
}

stable_periods <- Ciitizen_medication_stable %>%
 group_by(patient_uuid) %>%
 do(find_stable_periods(.))

stable_periods <- stable_periods %>% rename("UUID" = "patient_uuid")
study <- merge(study, subset(stable_periods, select = c(1, 4)), by = "UUID", all=T)
```

##3.2 Getting Seizure types 
```{r}
seizures <- Ciitizen_seizure 

seizures <- seizures %>% select(c(UUID, seizure_history_type))
seizures_onehot <- as.data.table(seizures)
seizures_onehot$seizure_history_type <- as.factor(seizures_onehot$seizure_history_type)
seizures_onehot <- one_hot(seizures_onehot)

seizures_onehot <- seizures_onehot %>%  group_by(UUID) %>% 
  summarise(across(c(`seizure_history_type_Absence seizure`:`seizure_history_type_Unprovoked seizure`), ~max(.x, na.rm = TRUE)))
study <- merge(study, seizures_onehot, by = "UUID", all=T)
```

##3.4 Getting Hospitalization Events
```{r}
hospitalization <- Ciitizen_hospitalization %>% select(patient_uuid, admission_diagnosis, admission_age_days_firstDate,
                                                       admission_age_days_lastDate)

hospitalization$admission_age_days_firstDate <- as.numeric(hospitalization$admission_age_days_firstDate)
hospitalization$admission_age_days_lastDate <- as.numeric(hospitalization$admission_age_days_lastDate)

length_stay <- apply(hospitalization[3:4],1, diff)
hospitalization <- cbind(hospitalization, length_stay)
hosp_onehot <- as.data.table(hospitalization)
hosp_onehot$admission_diagnosis <- as.factor(hosp_onehot$admission_diagnosis)
hosp_onehot <- one_hot(hosp_onehot)
hosp_final <- hosp_onehot %>% group_by(patient_uuid) %>% 
  summarise(across(`admission_diagnosis_Abnormal movement`:admission_diagnosis_Vomiting, sum, na.rm = T))
hosp_final <- hosp_final %>% rename("UUID"="patient_uuid")
study <- merge(study, hosp_final, by = "UUID", all = T)

hosp_stays <- hosp_onehot %>% group_by(patient_uuid) %>% 
  summarise(across(length_stay, mean, na.rm = T)) %>% rename("UUID"= "patient_uuid")

study <- merge(study, hosp_stays, by = "UUID", all = T)
```

##3.5 Demographics: Sex and Age at diagnosis
```{r}
dems <- Ciitizen_demographics %>% select(patient_uuid, sex, age_at_diagnosis) %>% 
  rename("UUID" = "patient_uuid")

#recode male to 1, female to 0
dems <- dems %>% mutate(sex = ifelse( sex == "male", 1,0))
dems$age_at_diagnosis[ dems$age_at_diagnosis== "undefined"] <- 100
dems$age_at_diagnosis <- as.numeric(dems$age_at_diagnosis)
dems <- dems %>% mutate(age_at_diagnosis_m = age_at_diagnosis*12)

study <- merge(study, dems, by = "UUID", all = T)
study <- study %>% mutate(onset_diagnosis_gap = ifelse(age_at_diagnosis_m<1200, (age_at_diagnosis_m - age_onset_m), "undefined")) %>% select(-age_at_diagnosis)
```

##3.6 Getting Clinical Diagnoses
```{r}
diagnosis <- Ciitizen_diagnosis %>% select(patient_uuid, clinical_diagnosis) %>% 
  rename("UUID" = "patient_uuid")

diagnosis <- as.data.table(diagnosis) 
diagnosis$clinical_diagnosis <- as.factor(diagnosis$clinical_diagnosis)
diagnosis <- one_hot(diagnosis)
diagnosis <- diagnosis %>% group_by(UUID) %>% 
  summarise(across(`clinical_diagnosis_Abdominal pain`:`clinical_diagnosis_West syndrome`, sum, na.rm=T))

study <- merge(study, diagnosis, by = "UUID", all =T)
```

##3.7 Diagnostic Procedure Results
```{r}
test_results <- Ciitizen_diagnostic %>% select(patient_uuid, procedure_findings) %>% 
  rename("UUID" = "patient_uuid")

test_results <- as.data.table(test_results)
test_results$procedure_findings <- as.factor(test_results$procedure_findings)
test_results_onehot <- one_hot(test_results)

test_results_final <- test_results_onehot %>% group_by(UUID) %>% 
  summarise(across(`procedure_findings_Abnormal restricted diffusion`:`procedure_findings_Videofluoroscopy swallow normal`, sum, na.rm=T))

study <- merge(study, test_results_final, by = "UUID", all = T)
write.csv(study, "./data/Supplementary_Table - Study Data by Groups.csv")
```
